
---
title: "Wikipediatrend Introduction"
author: 
  - Melissa Wilson
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
```

# Description

- There are many WikiMedia packages available with different purposes, but we will focus on Wikipediatrend
- Wikipediatrend is an interface between R and Wikipedia that allows you to quickly access page traffic to wiki pages


# Install

- First we will install the packages using `Install.packages("WikipediR", "WikipediaR", "Wikidata", "Wikipediatrend")`

- The only one we really need is wikipediatrend, but might as well get the other commonly used packages while you're at it


# Get Main Page Traffic Data

- We want to access daily page view counts
- The most important function you will need to know in order to use this package is `wp_trend()` because it will help us extract the page count data from Wikipedia

```{r eval=FALSE, echo=TRUE}
library(wikipediatrend)
page_views <- wp_trend("main_page")
```

- Note: You may see the message 
```{r}
"The wikipediatrend package currently has no server providing any page view information. Use package pageviews for recent (2016+) information. Older information hopefully available again soon."
```

- To get around this issue, I installed the development version

```{r eval=FALSE, echo=TRUE}
devtools::install_github("petermeissner/wikipediatrend")

library(wikipediatrend)
page_views <- wp_trend("main_page")
page_views
```

-Now we can see the data

```{r eval=FALSE, echo=TRUE}
ggplot(page_views, aes(x=date, y=views, group=article, color=article)) +
  geom_line(size=1) + theme_bw()
```


# Example 1

- Let's try some examples!

```{r eval=FALSE, echo=TRUE}
page_views1 <- 
  wp_trend( 
    page = c( "Fake_news")
  )
```

```{r eval=FALSE, echo=TRUE}
library(tidyverse)

ggplot(page_views1, aes(x=date, y=views, color = wp_day(date))) +
  geom_line(size=1) + 
  theme_bw()

```

- Or we can look at a specific time period

```{r eval=FALSE, echo=TRUE}
page_views2 <- 
  wp_trend( 
    page = c("Fake_news") ,
    from = "2018-04-16",
    to = "2018-04-30"
  )
```

```{r eval=FALSE, echo=TRUE}
library(tidyverse)

ggplot(page_views2, aes(x=date, y=views, color=wp_day(date))) +
  geom_line(size=1) + theme_bw()

```

- Notice the traffic drop on the weekends?


# Example 2

- What if you want to use multiple wiki articles?
- You simply list them!

```{r eval=FALSE, echo=TRUE}
page_views3 <- 
  wp_trend( 
    page = c( "Tropical_cyclone", "Saffir-Simpson_scale"), 
    from = "2016-01-01",
    to   = "2018-05-01" #prev_month_end() should also work, but it was not working for me 
  ) #Saffir-Simpson scale measures hurricane categories
```

- Be sure to specify the group as 'article' because you are using more than one page's traffic data

```{r eval=FALSE, echo=TRUE}
library(tidyverse)

ggplot(page_views3, aes(x=date, y=views, group=article, color=article, fill=article)) +
  geom_line() + theme_bw()
```
- We can see a giant spike about the time hurrican Irma turned up last year.


# Example 3

- Marvel page (No spoilers, I promise!)

```{r eval=FALSE, echo=TRUE}
page_views4 <- 
  wp_trend( 
    page = c( "List_of_films_based_on_Marvel_Comics"), 
    from = "2018-01-01",
    to = "2018-05-05"
  ) 
```

```{r eval=FALSE, echo=TRUE}
library(tidyverse)

ggplot(page_views4, aes(x=date, y=views, color=article)) +
  geom_line(size=1, color="pink") +
  theme_bw()
```

- There is an obvious spike around Apr 27th-29th due to the release of Avengers: Infinity War










